---
title: "Covid91 vaccine study Final2025"
author: "Israel Valdovinos"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
males <- FinalData %>% filter(sex == "M")
nrow(males)
male_test <- prop.test(sum(males$infected == 1), nrow(males), p = 0.5)
male_test
```
The test for males also produced a highly significant result, meaning that the vaccine has a statistically meaningful effect in this subgroup. Like with females, this indicates that the vaccine effectively reduces infection rates in male participants, providing evidence of its efficacy across sexes.

## Females

```{r}
females <- FinalData %>% filter(sex == "F")
nrow(females)
female_test <- prop.test(sum(females$infected == 1), nrow(females), p = 0.5)
female_test
```
The proportion test for females shows that the observed infection rate among female participants is significantly different from 50%. The very low p-value indicates strong evidence that the vaccine has a protective effect in this group. This suggests that the vaccine is effective for females, reducing the likelihood of infection beyond what would be expected by chance.

## LGBTQ

```{r}
lgbtq <- FinalData %>% filter(LGBTQ == "gay")
nrow(lgbtq)
lgbtq_test <- prop.test(sum(lgbtq$infected == 1), nrow(lgbtq), p = 0.5)
lgbtq_test
```

The analysis of LGBTQ participants shows a significant protective effect of the vaccine, with a p-value well below 0.05. This means that vaccination significantly lowers the risk of infection in this group, supporting the conclusion that the vaccine is effective across different sexual orientations.

## Druggies


# Overall Results and Conclusions